<center> 
 <font size="5">Changes in Weekly Wage & Weekly Working Hours by Industries</font>
</center> 

```{r, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center", warning = FALSE, message = FALSE)
```

```{r, echo=FALSE, message=FALSE}
#setwd("/Users/liwen/Documents/DV_final project/")
#getwd()
```

```{r, echo=FALSE, message=FALSE}
weekly_wage <- read.csv("wage4.csv")
```

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
```
```{r, echo=FALSE, message=FALSE}
library(plotly)
```

```{r, echo=FALSE, message=FALSE}
g1 <- ggplot(weekly_wage, aes(year, percentChange, text=paste("Time:",year,"<br />Industry:",industry,"<br />Percent change:",percentChange*100,"%"))) + scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + geom_path(aes(color = industry, group = industry)) + theme(axis.text.x = element_text(angle = 60, vjust = 0.5)) +  ylab("% Change Over the Year") +  theme(axis.title.x=element_blank(), panel.background = element_rect(fill='transparent') ) + ggtitle("Weekly wage (quarterly) % change over the year by industry") + theme(plot.title = element_text(size=12))
```
<center>  Weekly wages are one of the most important indicators that can reflect the state of the labor market. We aim to visualize the effect of the COVID-19 on weekly wage by different industries, within the time range of 2017 to 2021. Due to the large salary differences between various industries, we have selected the percentage change over the year for each industry. This provides the opportunity to observe fluctuations in weekly wage over time, as well as compare the trends among multiple industries by selecting the industries that interest you. This data was collected from the U.S. Bureau of Labor Statistics.
</center>

<center>
```{r, echo=FALSE, message=FALSE}
ggplotly(g1, tooltip = c("text")) 
```
</center>
<center>  The above visualization plot shows that the fluctuation pattern of weekly wage percentage change over the year varies in industries. At pre-Covid time, Weekly wages in all industries plummeted in the third quarter in 2017; at the beginning of the pandemic, arts, information and educational services industries had sudden increases, while other industries were either not affected too much, or had sharp declines (accommodation and mining); during the fourth quarter of 2020 and the first quarter of 2021, all the industries had sharp rises and sharp declines, respectively; some industries hit the lowest record of wage percentage change at the first quarter of 2021.    
 </center>
```{r, echo=FALSE, message=FALSE}
weekly_hours <- read.csv("hours4.csv")
```

```{r, echo=FALSE, message=FALSE}
hours_axis_var_names <- c("Construction", "Mining", "Manufacturing", "WholesaleTrade", "RetailTrade", "Transportation", "Information", "Finance", "RealEstate", "ProfessionalServices", "ManagementCompanies", "AdministrativeServices", "HealthCare", "Arts", "Accommodation")
```
 
```{r, echo=FALSE, message=FALSE}
create_hours_buttons <- function(weekly_hours, hours_axis_var_names) {
  lapply(
    hours_axis_var_names,
    FUN = function(var_name, weekly_hours) {
      button <- list(
        method = 'restyle',
        args = list('y', list(weekly_hours[,var_name])),
        label = sprintf(var_name)
      )
    },
    weekly_hours
  )
  
}
```

<br>
<br>
<center>  Working hours are another important indicator of the labor market. We aim to visualize how weekly working hours in different industries were affected by the COVID-19.  This data was collected from the U.S. Bureau of Labor Statistics.
 </center>
 <br>
```{r, echo=FALSE, message=FALSE}
 plot_ly(weekly_hours, x = ~time, y = ~Construction, type = 'scatter', mode = 'lines') %>%
    layout(
         title = "Average weekly hours by industry",
         xaxis = list(showgrid = FALSE, title = FALSE),
         yaxis = list(title = FALSE),
         updatemenus = list(
             list(
                 buttons = create_hours_buttons(weekly_hours, hours_axis_var_names),
                 pad = list('r'= 15, 't'= 10, 'b' = 0)
             )
         ))
```

<center>  The above visualization plot shows that workers in most industries have had to reduce their working hours as a result of the pandemic, the sharp reduction in working hours has been accompanied by an equally sharp decline in income that corresponds to the weekly wage visualization. At the same time, some industries such as arts, information and administrative services experienced increasing of working hours at the second quarter of 2020 (the initial stage of the Covid), which indicates the change of people's way of living -- reducing face-to-face activities and shifting to remote activities. 
 </center>
